\documentclass[10pt]{article}
\usepackage{hyperref,microtype,times,nth,textcomp,fontspec}
\usepackage[a4paper,left=.5in,top=0.5in,right=3.5in]{geometry}
% \usepackage{lipsum,color}

% \definecolor{Red}{HTML}{FF0000}
% \definecolor{Blue}{HTML}{0000FF}
% \newcommand{\ali}[1]{{\color{Red}[al2: #1]}}
% \newcommand{\topic}[1]{{\color{Blue}#1}} % if you're in a subfile, bold the \topic sentences to 

\usepackage[inline]{enumitem}
\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
\bibliography{references}
\AtBeginBibliography{\tiny}

\setmainfont{Source Serif Pro}[
 ItalicFont = Source Serif Pro,
 ItalicFeatures={FakeSlant=0.2}]
\setmonofont{Source Code Pro}

\pagenumbering{gobble}
\setlist[itemize]{noitemsep, nolistsep}
\setlist[enumerate]{noitemsep, nolistsep}

\begin{document}
% \marginpar{\lipsum[1]}

% \begin{enumerate}
  
\section{Some questions}
\begin{itemize}
  \item Is AI going to make our lives worse?
  \item Will AI help us understand people better, or will it be used to oppress people better?
  \item What circumstances characterize AI systems which \textit{service} people versus those that \textit{trade} in people?
\end{itemize}

\section{Approaches}
either or\dots
\begin{itemize}
  \item We present a framing for making sense of AI systems
  --- especially in the context of the above question ---
  through this lens that answers all these questions in useful ways.
  We also provide people with a useful foothold to begin ``climbing'' this otherwise unapproachable wall.
  \item We suggest a way of thinking about AI systems that may provide a useful perspective,
  specifically by aligning elements of this topic in such a way that
  certain answers become, or at least intuitive.
\end{itemize}


\section{Problem areas (and solution approaches?)}
\begin{itemize}
  \item advertising/surveillance state
  \begin{itemize}
    \item what happens? {why do i care?}
    \begin{enumerate}
      \item people exist (online)
      \item online systems make inferences about the people
      \item these systems put ads in front of people
    \end{enumerate}
  \end{itemize}
  \item justice/racism
  \begin{itemize}
    \item bureaucratic orgs do Y
  \end{itemize}
  \item streaming/garbage
  \begin{itemize}
    \item bureaucratic orgs do Z
  \end{itemize}
  \item ???
\end{itemize}

\newpage


  \section{hot question}
    \begin{itemize}
      \item Are AIs necessarily a dehumanizing force?
      \item Is the future of ``\textit{managed} work'' one of oppressive burdens?
      \item Are AIs going to be tools of humans, or will humans be tools of AIs? (extensions?)
      
      \item in what ways are AIs making decisions that affect us?
      \begin{itemize}
        \item legal: bail, sentencing, etc?
        \item financial: loans, insurance
        \item business: freelance work assignment, guiding workers to places
        \item politics: information diet
      \end{itemize}

    \end{itemize}
  
  \section{why does it matter?}
    \begin{itemize}
      \item The answer to (one of those questions) is going to determine if Neo from the Matrix is going to have to save us.
    \end{itemize}
  
  \section{prior work: what's missing?}
    \begin{itemize}
      \item \citeauthor{uberAlgorithm}~\cite{uberAlgorithm} looked at this, but didn't take a shot at the deeper issues
      \item \citeauthor{turkopticon}~\cite{turkopticon} got into some of the deeper undercurrents, but \dots
      \item \textsc{need to come up with other examples --- look at notes from CHI review; look at FATML papers}
      \item Amazon warehouse workers
      \item stitchfix? @mav
      \item Wikipedia?
      \item Facebook/Instagram/YouTubers working around the opaque preferences of the system?
      % \item Qbot, dan weld, etc\dots
      % \item 
    \end{itemize}
  
  \section{lay out the metaphor}
    \begin{enumerate}
      \item Bureaucracies are comprehensive organizations which try to achieve certain \textbf{goals}
      \item They achieve those \textbf{goals} by doing \textbf{tasks}
      \item The processes that determine how to do a task are what form the structure of bureaucracies
      \item Following the process is deeply important, but people within the bureaucracies
      --- bureaucrats ---
      are empowered to use \textit{discretion} in their own ways.
      \item This discretion is exercised at different levels and in different contexts to varying degrees, but
      always in service to the \textbf{goals} of the bureaucracy
      (e.g. doctors providing medical care, so misclassifying patients as inpatient)
    \end{enumerate}
    AIs are basically like that
    \begin{enumerate}
      \item Bureaucracies are the AIs
      \item They're embedded with certain goals
      \item They develop their own tasks to accomplish those goals (with incentives and disincentives to approach those goals)
      \item Adherence to the process is important
      \item \textsc{It's difficult to conceptualize AIs which can exercise discretion}
    \end{enumerate}
  
  \section{what does the metaphor tell us?}
    \begin{itemize}
      \item AIs lack the capacity to exercise discretion in the current design paradigm
      \item But this is necessary to deal with humans, because of the ``\dots~relatively high degree of uncertainty because of the complexity of subject matter (people) and the frequency or rapidity with which decisions have to be made.''
    \end{itemize}
  
  \section{how should we rethink algos?}
    \begin{itemize}
      \item We need to figure out how to design algorithms with mechanisms for people to intercede and break the rules on behalf of the higher order purpose
      \item AIs which have some conceptual understanding of that higher order purpose --- the \textbf{goals} of the org --- would be important
      \item Encoding the nuance of those \textbf{goals} will prove to be impossible
      \item Here are some ideas!
      \begin{itemize}
        \item Question the \textit{descriptive} vs \textit{prescriptive} tendencies of AI/ML
              (e.g. \textit{reflecting} societal prejudice vs implicitly \textit{encoding} it into the system)
        \item Provide operators with well--documented vectors to mess with levers
        \item Do the things from that ethics paper about making clear how a decision was made
      \end{itemize}
    \end{itemize}
% \end{enumerate}


\section{Refs}
\printbibliography[heading=none]
% \printbibliography{}
\end{document}