\documentclass[main]{subfiles}

% \onlyinsubfile{
%   \usepackage[T1]{fontenc}
%   \setsansfont{Source Sans Pro}
% }

\begin{document}

\section{Introduction}
\subsection{review some problems in AI systems}
\subsection{suggest Theoretical basis for thinking about this stuff}
\subsection{Enter \texttt{theory}}
\section{Review of \texttt{theory}}
\section{Theories}

\subsection{Idea 1: Bureaucratic Theory}
\subsubsection{Issue}
AI systems learn from societal data that reflects encoded bias, resulting in biased AI.

\subsubsection{How can we think about this using bureaucratic theory?}
AI systems get commissioned to solve a high level problem --- a goal.
The skills and perspectives (and shortcomings) of the people who set out to build that AI encode their biases in similar ways to bureaucracies.
Those biases get baked into the system in a way that becomes difficult to notice except with careful reflection.

\subsubsection{Methodological wins?}

\subsubsection{What would this perspective suggest? What questions would we still need to answer?}
???



\subsection{Idea 2: We're all living under Feudalism}
\subsubsection{Issue}
People do work for platforms with little to no say in how the systems are operated.
We benefit superficially, but don't meaningfully own anything.

\subsubsection{How can we think about this using feudalism?}
People do work for the benefit of online platforms that generates value the people never see.
They don't own any part of the system, or the ``ground'' under which they work, and they have no input into the politics of the system.
If Flickr gets sold, the people living on the land just have a new lord.

\subsubsection{Methodological wins?} Economic anthropology, \citeauthor{polanyi1944great}, etc\dots become available ways to think about societal transformations that empower people's relationships with the world.

\subsubsection{What would this perspective suggest? What questions would we still need to answer?}
???

keep an eye on the interaction between the users and (algorithms, designers, ???)

search bias (see downloads)


(how people deal with AIs)

latverian AI


\subsection{Idea 3: ???}
\subsubsection{Issue}

\subsubsection{How can we think about this using blargh?}
\subsubsection{Methodological wins?}

\subsubsection{What would this perspective suggest? What questions would we still need to answer?}

\section{Discussion}
\section{Conclusion}



\onlyinsubfile{
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{references}
}
\end{document}
